---
title: Variable importance in ANNs
output: html_document
---

```{r setup, include=FALSE}
source(file.path(PROJHOME, "src/r/setup.R"))
setup()
```

Examples are from the following:

> Ibrahim, OM. 2013. [A comparison of methods  for  assessing the  relative  importance of input variables in  artificial neural networks](http://www.palisade.com/downloads/pdf/academic/DTSpaper110915.pdf). Journal of Applied Sciences Research, 9(11): 5692-5700.

Table 4: Final connection weights
```{r network-weights}
w <- NULL
w$ih <-
  data.frame(x1 = c(-0.8604, 1.0651, -0.1392, 1.0666),
             x2 = c(-6.6167, 1.5142, -5.3051, 1.3358),
             x3 = c(-1.7338, 1.4107, -1.4207, 0.7845)) %>%
  t() %>%
  `colnames<-`(c("h1", "h2", "h3","h4"))

w$ho <-
  data.frame(gy = c(-5.2075, 2.0158, -4.4617, 1.5858)) %>%
  t() %>%
  `colnames<-`(c("h1", "h2", "h3","h4"))

w
```

<br>
<br>

### Connection weights algorithm (CW)

$RI_x = \sum_{y=1}^m{w_{xy}w_{yz}}$

----

Here we show explicitly the calculations for input `x1`

```{r echo=TRUE}
# We first calculate the connection weight product of eg. input x1
# w_x1,h1 * w_h1,gy
# w_x1,h2 * w_h2,gy
# w_x1,h3 * w_h3,gy
# w_x1,h4 * w_h4,gy
(cw_prd_x1 <- w$ih[1,] * w$ho)

# & then take the sum
sum(cw_prd_x1)
```

----

We do the calculation for all inputs and get the following connection weight product sums (`cw_prd_sums`):

```{r}
(cw_prd_sums <-
  sapply(1:3, function(i){sum(w$ih[i,] * w$ho)}) %>%
  setNames(., rownames(w$ih))
 )
```

To calculate relative importance, we normalize and sort to find the most important feature

```{r echo=TRUE}
(cw_prd_sums / sum(cw_prd_sums)*100) %>%
  sort(decreasing = TRUE)
```
