---
title: "Features"
---

```{r setup, include=FALSE}
source(file.path(PROJHOME, "src/r/setup.R"))
setup()

library(rtracklayer)
```

```{r define-files}
gene_file <- file.path(PROJHOME, "reference/Homo_sapiens.GRCh37.75.gtf.chr.pcgene")
bin_file <- file.path(PROJHOME, "reference/hg19.200bpbins.bed.srt")

promoter_file <- file.path(PROJHOME, "results/features/promoter",
                            "Homo_sapiens.GRCh37.75.gtf.chr.pcgene.promoters.bed")
binfeature_file <- file.path(PROJHOME,  "results/features/promoter/bins-around-promoter.bed")
```

## Preface

- 3 trillion base pairs (bp) in human genome
- dividing the genome into 200 bp bins produces 15 million bins
- 15 million features is a lot
- thus will focus on the bins around promoters of protein coding genes

## Identify features

```{r load-pcgenes}
gr <-
  import.gff(gene_file) %>%
  keepStandardChromosomes()

gr.promoters <-
  promoters(gr, upstream = 0, downstream = 1)

# save data
data.frame(
  chr = seqnames(gr.promoters),
  start = start(gr.promoters),
  end = end(gr.promoters)
  #gene_id = mcols(gr.promoters)$gene_id
  ) %>%
  tbl_df() %>%
  readr::write_tsv(promoter_file, col_names = FALSE)
```

- Protein coding genes are loaded from `r basename(gene_file)`
- `r length(gr.promoters)` protein coding genes on the standard chromosome

```{r cmd-identify-features, results='asis'}
# The 11 bins around a TSS
cat("```\n",
    sprintf("bedtools window -a %s -b %s -w 1000 | cut -f4-6 | sort | uniq | bedtools sort -i - > %s",
            promoter_file, bin_file, binfeature_file),
    "\n```")
```

```{r count-features}
num_features <- NULL
if (file.exists(binfeature_file)) {
  num_features <-
    rtracklayer::import.bed(binfeature_file) %>%
    length()
}
```

- `r num_features` unique bins (features) span the 11 bins around the
  `r length(gr.promoters)` promoters

## Extract features (example)

```{r extract-features-example-files}
finder_file <- "/projects/edcc_new/reference_epigenomes/CEMT_19/bams/ChIP-Seq/H3K4me3/FindER.1.0.0b/A33074.H3K4me3.merge_bwa.hg19a.A33074_2_lanes_dupsFlagged.v.A33079.Input.FDR_0.05.FindER.bed.gz"
samplefeature_file <- "/projects/csiu_prj_results/PROJECTS/predictTissue/results/features/promoter/test-sample.EF.bed"
sample_file <- "CEMT_19"
```

```{r cmd-extract-features, results='asis'}
cmd_extract_features <-
      sprintf("bedtools intersect -a %s -b %s -c",
              binfeature_file, finder_file)
cmd_parser <-
  sprintf("awk -F'\\t' -v sample_id='%s' 'BEGIN{print sample_id} {print $NF}'",
          sample_file)
cmd_writeout <-
  sprintf("> %s", samplefeature_file)

cat("```\n",
    paste(cmd_extract_features, "|", cmd_parser, cmd_writeout),
    "\n```")
```

Then can combine features with `paste file1 file2 file3`
